---
title: "HW04 Part 1: Complete Chapter 3"
author: "Ryan McAnulty"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 3

- Change "your name" in the YAML header above to your name.

### Section 3.1

Load the tidyverse packages. I've entered the first code chunk for you. **You must enter all subsequent code chunks and run the code.** 

```{r}
library("tidyverse")
```

If you get an error, then you did not install the `tidyverse` package like you were supposed to in the previous assignment. Go back to that assignment and complete that step. (This should not be the case, though.)

### Section 3.2: First steps

Enter your code chunks for Section 3.2 here.
### 3.2.2 Creating a ggplot

## A plot of mileage of several car models displaying engine displacement vs. highway fuel economy
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

Get in the habit of entering a brief description of what each chunk of code does. Enter this above the chunk, like I did in Section 3.1.

#### Section 3.2 Questions

Answer the questions *completely.* Some answers will require you to write code in a code chunk and also to type text before or after the chunk. Other answers will require only code or only text. 

**1:** Run ggplot(data = mpg). What do you see?
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```
Tis displays a scatterplot of mpg with engine displacement vs. hwy fuel economy


**2:** How many rows are in `mpg`? How many columns? **Hint:** Use the `dim()` function. Type `dim()` in the console to learn what `dim()` does.

## Displays dimensions of mpg
```{r}
dim(mpg)
```
Mpg has 234 rows and 11 columns.

**3:** What does the `drv` variable describe? Read the help for `?mpg` to find out.
The drv variable refers to the type of drive train of a vehicle: f = front-wheel drive, r = rear-wheel drive, 4 = 4wd

**4:** Make a scatterplot of `hwy` vs `cyl`.
## A plot of hwy vs. cyl
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```


**5:** What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?
## A plot of class vs. drv
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```
The plot is not useful because the variables are not a function of each other, a table or some other way of conveyin information
would be better for these variables.

### Section 3.3: Aesthetic mappings
## A scatterplot of displ vs. hwy also showing different classes as different colors
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x= displ, y = hwy, color = class))
```
## Same plot as above, but the classes are denoted by a different sized data point and are all black

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y =hwy, size = class))
```
## Again, displ vs. hwy with the aesthetic variable alpha used for class type
```{r}
 ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y =hwy, alpha = class))
```
## Plot of displ vs. hwy with the shape aesthetic used for class
```{r}
 ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y =hwy, shape = class))
```

## Plot of displ vs. hwy with the color of the data points manually set to blue
```{r}
 ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y =hwy), color = "blue")
```


#### Section 3.3 questions

**1:** What’s gone wrong with this code? Why are the points not blue?
## Plot of displ vs. hwy with the intention of the data points being
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```   
To manually set an aesthetic value, it has to be outside of the first set of parentheses.  Inside the parentheses, ggplot sets the third variable as a legend and in this case, named blue/

**2:** Which variables in `mpg` are categorical? Which variables are continuous? (*Hint:* type `?mpg` to read the documentation for the dataset). How can you see this information when you run mpg?

The categorical variables are "manufacturer", "model", "trans", "drv", "fl", and "class".
The continuous variables are "displ", "year", "cyl", "cty", and "hwy".  
When you run mpg, their is a label right under the column name.  <chr> represents categorical variables, just characters, not numeric.  <dbl> and <int> denotes continuous variables like numerals.

**3:** Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables? **Note:** You need several code chunks here to answer this question.

## disp vs. hwy using color with a continuous variable
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = year))
```
Using color with a continuous variable displays a the same color, but changes the hues from bright to dark.

## Color using a categorical variable.
```{r}
 ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y =hwy, color = drv))
```
Using color with a categorical variable gives different distinct colors for each variable.

##  Using the aesthetic size with a continous variable
```{r}
 ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y =hwy, size = cty))
```
This rounds the variable to the nearest interval and the intervals span a range that the program sees as representative of the values.

## Shape with a categorical variable
```{r}
 ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y =hwy, size = trans))
```
The plot works fine with a different size for each variable, but a message is displayed that says "Using size for a discrete variable is not advised."

## Using the Shape aesthetic with a continuous variable
```{r}
 ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y =hwy, shape = cyl))
```

This will not plot : Error: A continuous variable can not be mapped to shape.

## Using shape with a categorical variable
```{r}
 ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y =hwy, shape = manufacturer))
```
This will plot the first six values with different shapes but displays no shapes for the remaining ones.  An error message appears that says I would have to set each variable shape manually, even then, there are only 25 different shapes in R.

**4:** What happens if you map the same variable to multiple aesthetics?
## Setting year to both color and size
```{r}
 ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y =hwy, color = year, size = year))
```
It plots with a range to the color hue and size.  Shape doesn't plot with continuous.

## Using a categorical variable with size and color
```{r}
 ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y =hwy, color = drv, size = drv))
```
This assigns both a size and color to categorical variables.

**5:** What does the stroke aesthetic do? What shapes does it work with? (*Hint:* use `?geom_point`)

## Using the stroke aesthetic and manual shape
```{r}
 ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y =hwy, color = cyl, stroke = 3), shape = 13)
```

Stroke seems to be the "thickness" of the data point, like a thicker paint brush or pencil lead thickness the higher the stroke value.  It seems to work with all the shapes.

**6:** What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`?
## Using variable name other than a variable name
```{r}
 ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y =hwy, color = displ < 5, stroke = 1))
```

Using a function like this, will display a logic plot with two variables, TRUE(if the variable value satisfies the function) or False (if it does not).

### Section 3.5: Facets

## Using facet wrap to show displ vs. hwy for all classes in mpg in 2 rows
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2)
```

## Using facet grid of displ vs. hwy along with drv and cyl
```{r}
ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y = hwy)) +
   facet_grid(drv ~ cyl)
```

#### Section 3.5 questions

**1:** What happens if you facet on a continuous variable? **Note:** Write a prediction, then make and run your code chunk.
Prediction: It will produce a gradient similar to previous plots
```{r}
ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y = hwy)) +
   facet_wrap(~ cty, nrow = 2)
```
I was wrong, it plotted the same


**2:** What do the empty cells in plot with `facet_grid(drv ~ cyl)` mean? How do they relate to this plot?
That there are no values to place in those cells.  In this case, there are no vehicles that have 5 cylinders AND are 4 wheel drive or rear wheel drive or a 4 cylinder that is rear wheel drive.


**3:** What plots does the following code make? What does `.` do?
## Facet grid with displ vs. hwy and drv in rows and . in columns
```{r}
ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, hwy)) +
   facet_grid(drv ~ .)
```
This makes a plot without a facet in the columns

## Facet grid with displ vs. hwy and . in rows and cyl in columns
```{r}
ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, hwy)) +
   facet_grid(. ~ cyl)
```
This makes the plot without a facet in the rows

**4:** Take the first faceted plot below. What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
```{r}
ggplot(dat = mpg) +
   geom_point(mapping = aes(x = displ, y = hwy)) +
   facet_wrap(~ class, nrow = 2)
```
Some advantages are that the plots are less cluttered, if data sets overlap, you can't see that.  Shows in a specific category.
Disadvantages are that it is a little harder to compare to different data points, takes up more room.
When the there are more data points, the advantages of separation and more space of faceting would become more pronounced.

**5:** Read `?facet_wrap`. What does `nrow` do? What does `ncol` do? What other options control the layout of the individual panels? Why doesn’t `facet_grid()` have nrow and ncol argument?
?facet_wrap

nrow sets the number of rows.
ncol sets the number of columns.
Other controlers are scales, shrink, labeller, as.table, switch, drop, dir, and strip.position.
The number of rows and columns in facet_grid is determined by the number of different values of the data points.  For example year has 1999 and 2008, so it would have 2 rows or columns. 

**6:** When using `facet_grid()` you should usually put the variable with more unique levels in the columns. Why?

To take up less vertical space in the plot; the y-axis would be "smooshed" and be reader to read
.
### Section 3.6: Geometric objects
## Scatter plot of displ vs. hwy using geom_point
```{r}
ggplot(data = mpg) +
   geom_point(mapping = aes(x = displ, y = hwy))
```
## Trend line plot of displ vs. hwy using geom_smooth
```{r}
ggplot(data = mpg) +
   geom_smooth(mapping = aes(x = displ, y = hwy))
```

## Using linetype to display a third variable
```{r}
ggplot(data = mpg) +
   geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```

## Trendline plot of displ vs. hwy
```{r}
ggplot(data = mpg) +
   geom_smooth(mapping = aes(x = displ, y = hwy))
```

## Using group to leave out the legend for drv
```{r}
ggplot(data = mpg) +
   geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
```

## Using show.legend = FALSE to leave out legend, but showing different colors of drv
```{r}
ggplot(data = mpg) +
   geom_smooth(
      mapping = aes(x = displ, y = hwy, color = drv),
      show.legend = FALSE
   )
```

## Showing both scatter plot and trendline plot for displ vs hwy in same plot
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

## Same plot, but with less code
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
   geom_point() +
   geom_smooth()
```

## Same plot, but with color asethetic added to scatter plot only
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
   geom_point(mapping = aes(color = class)) +
   geom_smooth()
```

## Using filter to single out the trend line for subcompact cars
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
   geom_point(mapping = aes(color = class)) +
   geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

#### Section 3.6.1 Questions

**1:** What geom would you use to draw a line chart? A boxplot? A histogram? An area chart? **Note:** You will have to make some types of these charts in Part 2 of this assignment.
I would use geom_line for a line chart.
geom_boxplot for a boxplot.
geom_freqpoly for a histogram
geom_ribbon for are an area chart


**2:** Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.
It will display a chart with scatter points for displ and hwy with colored drv points and a legend for drv and also 3 trend lines but without a confidence interval displayed

## Displaying discussed plot
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
   geom_point() +
   geom_smooth(se = FALSE)
```

I was correct

**3:** What does `show.legend = FALSE` do? What happens if you remove it?
**Note:** Skip the "Why do you think I used it earlier in the chapter?" question.
Show.legend = FALSE tells the program to not display the legend.
Leaving it out defaults to show.legend = TRUE and a legend displays.

**4:** What does the `se` argument to `geom_smooth()` do?
It controls whether or not the confidence interval is shown around the trend line.


**5:** Will these two graphs look different? Why/why not?
No, they should do the same.  The first sets up the proper info in the ggplot line.  The second waits to do it in each geom.  The first is simply less typing/code.

**6:** Recreate the R code necessary to generate the following graphs. **Note:** Enter a separate code chunk for each graph, numbered 1 to 6, starting row-wise at upper left. 

1: upper left.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
   geom_point() +
   geom_smooth(se = FALSE)
```

2: upper right.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group = drv)) +
   geom_point() +
   geom_smooth(se = FALSE)
```


3: middle left.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group = drv, color = drv)) +
   geom_point() +
   geom_smooth(se = FALSE)
```

4: middle right.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
   geom_point(mapping = aes(color = drv)) +
   geom_smooth(se = FALSE)
```

5: lower left.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group = drv)) +
   geom_point(mapping = aes(color = drv)) +
   geom_smooth(mapping = aes(linetype = drv), se = FALSE)
```


6: lower right. *Hint:* Read the help file for `geom_point`. Use `shape = 21`, `stroke = 2`, and `size = 3`.
```{r}
ggplot(data = mpg) + 
   geom_point(mapping = aes(x = displ, y = hwy, group = drv, fill = drv), color = "white", shape = 21, stroke = 2, size = 3)
```

### Section 3.7: Statistical transformations.

**Skip this section.** We may return to it later.


### Section 3.8 Position adjustments

## Displays bar chart of diamonds with color
```{r}
ggplot(data = diamonds) +
   geom_bar(mapping = aes(x = cut, color = cut))

```

## Displays a bar chart of diamonds with fill
```{r}
ggplot(data = diamonds) +
   geom_bar(mapping = aes(x = cut, fill = cut))
```

## Displays a bar chart of diamonds stacking cut and clarity

```{r}
ggplot(data = diamonds) +
   geom_bar(mapping = aes(x = cut, fill = clarity))
```

## Displaying plot with identity to remove stacking with transluecent bars
```{r}
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = 1/5, position = "identity")
```

## Plot with identity to remove stacking with no fill
```{r}
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) + 
  geom_bar(fill = NA, position = "identity")
```

## Plot using position set to fill
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```

## Using dodge to avoid stacking issue
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```

## Using jitter to avoid overplotting
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")
```

#### Section 3.8 Questions
**1:** What is the problem with this plot? How could you improve it. *Hint:* Remember the size of the `mpg` data set from earlier in this exercise. Does the number of points seem correct? **Note:** Write the code to improve it. 
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
   geom_point()
```
It is simply a plot of city fuel economy and highway fuel economy with overlapping and no real context.  This plot, as is, is not really saying anything.
To improve, I would use "jitter" to prevent over lap.  Add color or shapes to the different classes or drive trains.  Perhaps add a treadline to show overall trend.   

**2:** What parameters to `geom_jitter()` control the amount of jittering?
height and width control the amount of spread of the jitter. 

**3:** Compare and contracts `geom_jitter()` with `geom_count()`.

Both deal with solving Overplotting, but in different ways 
Geom_jitter introduces a little noise of a random amount to "stutter" the data points to get rid of points overlapping.
Geom_count displays all points as is, but counts how many points, if any, are occupying the same space.  It then displays how many actual points are at that one point.

**4:** What is the default position adjustement for `geom_boxplot()`? Create a visualization of the `mpg` dataset that demonstrates it.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_boxplot()
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_boxplot(mapping = aes(position = position_dodge2(width = 0.8)))
```


### Section 3.9: Coordinate systems

**Note:** Read about and do only the code chunk associated with `coord_flip()`. `coord_flip()` is a great layer to use when you have categorical data with longish names as it makes the text easier to read.

## Box plot of mpg class vs. hwy
```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
```

## Same but using coord_flip to flip x and y axes
```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```

### Section 3.10: The layered grammar of graphics

Just read this section as a contextual review of everything you just learned.


