---
title:"Useful Title"
author:"Your Name"
date:"`r format(Sys.time(),'%d %B %Y')`"
output:html_notebook
editor_options:
  chunk_output_type:inline
---
## HW 03
## Part 1: An R Notebook Case Study

## Loads tidyverse, kitr, and xtable
```{r}
library(tidyverse)
library(knitr)
library(xtable)
```

## Displays anscombe
```{r}
anscombe
```

## Creates dataset ansc
```{r}
ansc <- 
  bind_cols(
    anscombe %>%
      select(matches("x")) %>%
      gather(dataset, "x"),
    anscombe %>%
      select(matches("y")) %>%
      gather(key, "y")
  ) %>%
  select(-key) %>%
  mutate(dataset = gsub("x","", dataset)) %>%
  as_tibble()
```

## Displays dataset ansc
```{r}
ansc
```

## Runs and displays a summary data set of a statistical analysis of ansc datasest
```{r}
ansc %>% 
  group_by(dataset) %>% 
  summarize(
    mean_x = mean(x),
    stdev_x = sd(x),
    mean_y = mean(y),
    stdev_y = sd(y)
  )
```

## Creates and displays plots of ansc at two plots per column
```{r}
ansc %>% ggplot(aes(x,y, color = dataset)) +
  geom_point() +
facet_wrap(~dataset, ncol = 2)
```

## Part 2: Two more R Markdown tricks

## Displays frist six rows and header column names of dataset iris
```{r}
head(iris)
```

## Create table of iris head, renaming column names and aligning columns
```{r}
kable(head(iris),
      align= c("l", "c", "r", "c","l"),
      col.names = c("Sepal Length", "Sepal Width", "Sepal Length", "Sepal Width", "Species"))
```

## Display head of dataset airquality
```{r}
head(airquality)
```

## Create a table of head of airquality, renaming column names and aligning columns
```{r}
kable(head(airquality),
      align= c("r", "r", "c", "c","l","l"),
      col.names = c("Ozone (ppb)", "Solar R (lang)", "Wind (mph)", "Temp (°F)", "Month", "Day"))
```

## Create value chicks_lm from dataset  chickwts and running a summary of it
```{r}
chicks_lm <- lm(weight ~ feed, data = chickwts)

summary(chicks_lm)
```

## Attempt to form data set of chicks_lm.  Displays Error in as.data.frame.default(x)
## Cannot coerce class '"lm"' to a data frame.  If you try to do just the head from chickwts, it spits out
## values.  I speculate that kable() runs the data the isn't condusive with the lm function.
```{r}
kable(chicks_lm)
```

## Running chicks_lm with xtable
```{r}
xtable(chicks_lm) %>% kable()
```

## Run xtable and kable on trees data set to determine if Girth is a function of Height
```{r}
trees_lm <- lm(Girth ~ Height, data = trees)

xtable((trees_lm)) %>% kable()
```

## Bernoulli's Equation using LaTeX
$$p = \frac{\1}{\2}ρ + ρgh = C$$

^[This is a footnote]

