---
title: "HW05 Part 1: Data Visualization"
author: "Ryan McAnulty"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading tidyverse packages
```{r}
library(tidyverse)
```

#### Plot 1
## Scatter plot for dataset trees
## loading dataset "trees"
```{r}
data(trees)
```

## Checking format of data
```{r}
?trees
```

## Scatter plot of Height vs. Girth using dataset trees.  This shows the height (in feet) of 31 felled black cherry trees vs. the diameter (in inches) of the trees.
```{r}
ggplot(data = trees) +
  geom_point(mapping = aes(x = Height, y = Girth))
```
There is an over all trend that as the trees get taller, they also increase in diameter.

#### Plot 2
## loading dataset "lynx"
```{r}
data(lynx)
```

## Checking format of data
```{r}
?lynx
```

## Making a vector of the lynx data from the years 1821 to 1934
```{r}
year <- c(1821:1934)
```

## Detrmining what class dataset lynx is
```{r}
class(lynx)
```

## Converting lynx data to vector pelts
```{r}
pelts <- as.vector(lynx)
```

## Creating data frame lynx_pelts out of vectors year and pelts
```{r}
lynx_pelts <- data.frame("year" = year, "pelts" = pelts)
```

## Creating Linegraph plot of number of pelts gathered in Canada per year from 1821-1934
```{r}
ggplot(data = lynx_pelts) +
  geom_line(mapping = aes(x = year, y = pelts), color = "maroon")
```

It seems that the number of pelts gathered during these provided years followed an up and down trend.  There's a spike in the amount of pelts every 10 years or so and goes down to nearly 0 after 5 years.

#### Plot 3
