---
title: "HW05 Part 1: Data Visualization"
author: "Ryan McAnulty"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading tidyverse packages
```{r}
library(tidyverse)
```

#### Plot 1
## Scatter plot for dataset trees
## loading dataset "trees"
```{r}
data(trees)
```

## Checking format of data
```{r}
?trees
```

## Scatter plot of Height vs. Girth using dataset trees.  This shows the height (in feet) of 31 felled black cherry trees vs. the diameter (in inches) of the trees.
```{r}
ggplot(data = trees) +
  geom_point(mapping = aes(x = Height, y = Girth))
```
There is an over all trend that as the trees get taller, they also increase in diameter.

#### Plot 2
## loading dataset "lynx"
```{r}
data(lynx)
```

## Checking format of data for lynx
```{r}
?lynx
```

## Making a vector of the lynx data from the years 1821 to 1934
```{r}
year <- c(1821:1934)
```

## Detrmining what class dataset lynx is
```{r}
class(lynx)
```

## Converting lynx data to vector pelts
```{r}
pelts <- as.vector(lynx)
```

## Creating data frame lynx_pelts out of vectors year and pelts
```{r}
lynx_pelts <- data.frame("year" = year, "pelts" = pelts)
```

## Creating Linegraph plot of number of pelts gathered in Canada per year from 1821-1934
```{r}
ggplot(data = lynx_pelts) +
  geom_line(mapping = aes(x = year, y = pelts), color = "maroon")
```

It seems that the number of pelts gathered during these provided years followed an up and down trend.  There's a spike in the amount of pelts every 10 years or so and goes down to nearly 0 after 5 years.

#### Plot 3
## Loading dataset "iris"
```{r}
data(iris)
```

## Checking fromat of data for iris
```{r}
?iris
```

## Plotting scatterplot with iris data set Petal Length vs. Petal Width with distinct shapes and colors for species point size 2.
## Using labs to remove the "." out of the axes labels
```{r}
ggplot(data = iris) +
  geom_point(mapping = aes(x = Petal.Length, y = Petal.Width, color = Species, shape = Species), size = 2) +
  labs(x = "Petal Length", y = "Petal Width")
```
The Species "setosa" has much smaller petals than the other two.  "Virginica" is generally larger than "Versicolor" with a couple of the "Versicolor" being larger than the smallest of the "Virginica".

#### Plot 4 and 5
## Loading dataset msleep
```{r}
data(msleep)
```

## Checking data msleep format
```{r}
?msleep
```

## Looking up infor on plot type "Violin
```{r}
?geom_violin
```

## A Violin plot of dataset msleep with vore (diet of mammal) vs."rem_sleep" (Number of hours of REM sleep in hours) with fill color of 
## "gray30" 
```{r}
ggplot(data = msleep) +
  geom_violin(mapping = aes(x = vore, y = sleep_rem), fill = "gray30")
```
## A Violin plot of dataset msleep with vore (diet of mammal) vs."rem_sleep" (Number of hours of REM sleep in hours) with fill color of 
## "gray 70"
```{r}
ggplot(data = msleep) +
  geom_violin(mapping = aes(x = vore, y = sleep_rem), fill = "gray70")
```
It looks like meat eaters have the largest range of rem sleep hours per night, but usually requires about 2 hours.  It appears those that are plant eaters in general require less rem, perhaps so that they can recharge and avoid preadators more often.
It seems that geom_violin displays the data in a density plot, one that is wider where more data points converge.  It produces a plot that looks similar to the shape of a violin (to varying success).  "grey" displays a medium grey tint, added a number changes to a range of tint; from "grey 0" being nearly black to "grey 100" being basically white.

#### Plot 7
## Box plot of msleep dataset with order vs. sleep_total with the x and y axes flipped
```{r}
ggplot(data = msleep) +
  geom_boxplot(mapping = aes(x = order, y = sleep_total)) +
  coord_flip()
```
It seems that "rodentias" have has the second most range in total sleep, on average sleeping half the day. "artiocdactyla" sleeps very little,except for one outlying species that sleeps for 9 hours.  "chiroptera" sleeps almost all day.

#### Plot 8
## Box plot with points of msleep Conservation vs. Awake with the points "jittered" and Conservation colored
## Capitalizing the axes and Legend titles
```{r}
ggplot(data = msleep, mapping = aes(x = conservation, y = awake, color = conservation)) +
  geom_boxplot() +
  geom_point(position = "jitter") +
  labs(x = "Conservation", y = "Awake", color = "Conservation")
```
I'm honestly not really seeing any real trend of hours being awake depending on conservation status.  Perhaps a different plot geom would show a trend.

#### Plot 9
