---
title: "HW 07: Data Importing and Wrangling 1"
author: "Ryan McAnulty"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

#### Limpets
## Loading tidyverse and reading limpets.csv data
```{r}
library(tidyverse)
library(here)
```

The data is tidy.

## Defining limpets_data and forcing column types
```{r}
limpets_data <- read_csv(here("data", "limpets.csv"), col_types = "ncc")
```

## Setting levels for Site 
```{r}
limpets_data$Site <- factor(limpets_data$Site, levels = c("PBL", "KNRM", "VBG", "WP", "PF", "DP", "SIO", "CTZ1", "CTZ2", "CNMM"), ordered = TRUE)
```

## Box plot of limpets_data Length (mm) vs. Site with Protected status colored
```{r}
ggplot(data = limpets_data) +
  geom_boxplot(mapping = aes(x = Length, y = Site, color = Protected)) +
  labs(x = "Length (mm)")
```
The DP and WP sites have some outliers longer than the rest of the data.  For the most part, the protected sites tend to have larger Limpets.

#### Roseate Terns
## Reading and Naming data terns
```{r}
terns <- read_tsv(here("data", "roseate_terns.txt"))
```

## Viewing data
```{r}
terns
```

## Tidying data
```{r}
terns <- terns %>% 
  pivot_longer(c(AR:JC), names_to="Site",values_to="Population")
```

## Confirming tidying
```{r}
terns
```

## Filtering out NAs
```{r}
terns_complete_data <- filter(drop_na(terns))
```

## Confirming NAs removed
```{r}
terns_complete_data
```

## Line graph of Population vs. year of each site
```{r}
ggplot(data = terns_complete_data)+
  geom_line(mapping=aes(x=Population,y=year,color=Site))+
  labs(y="Year")
```
AC, AR, BC, JC obviously increase in population over time.

BR, JR had obvious decreases in population over time.

The line breaks come from haveing no data from that site in that year.

#### Blacklip Abalone
## Chunk 1: Reading abalone data, removing first column, and making a boxplot of each type
```{r}
abalone <- read_csv(here("data", "abalone.csv"))
abalone_data <- select(abalone, -(X1))
ggplot(data = abalone_data) +
  geom_boxplot(mapping = aes(x = Height, y = Type))
```

## Chunk 2: Scatterplot of Height vs. Rings to determine if outliers are old
```{r}
ggplot(data = abalone_data) +
  geom_point(mapping = aes(x = Height, y = Rings, color = Type))
```
No, it does not appear that the two outliers are exceptionally old, they both have <= 10 rings.

## Chunk 3: Filtering out the two outliers as coding errors, Ordering Type from I to M to F.  Replotting to see trends.
```{r}
abalone_cleaned <- filter(abalone_data, Height <= 0.3)
factor(abalone_cleaned$Type, levels = c("I", "M", "F"), ordered = TRUE)
ggplot(data = abalone_cleaned) +
  geom_point(mapping = aes(x = Height, y = Rings, color = Type))
```
For the most part, the Immature ones are the shortest and have fewer rings; they are juveniles.  The males and females seem to be evenly dispersed with no real trends on height or rings bases on sex.

#### Chunk 4: Assigned NA to Height = 0 and plotting results
```{r}
abalone_cleaned[abalone_cleaned$Height == 0, 'Height'] <- NA
ggplot(data = abalone_cleaned) + 
  geom_point(mapping = aes(x = Height, y = Rings, color = Type))
```

#### Chunk 5:  Scatter plots of Whole Weight vs. Age and Shucked Weight vs. Age of all three types of abalone.
```{r}
ggplot(data = abalone_cleaned, mapping = aes(x = WholeWeight, y = Rings, color = Type, shape = Type)) +
  geom_point() +
  geom_jitter()

ggplot(data = abalone_cleaned, mapping = aes(x = ShuckedWeight, y = Rings, color = Type, shape = Type))+
  geom_point() +
  geom_jitter()
```

As the abalone age, they get heavier until they have about 7 or so Rings, then the weight doesn't really seem to correlate with age.
Immature abalone tend to stay below 1 pound, Males and females vary.

Again, the shucked weight vs. age follows very closely to thee whole weight vs. age.  I would think that the older the albalone, the more meat it would yeild, there shucked weight follows age a little more than wholeweight, but not nearly as much as I would have thought. 

#### Darters
## Importing Darters data.  Setting Columns, naming coloumns, and setting orders
```{r}
darter_data <- read_fwf(here("data", "darter.txt"), fwf_widths(c(3, 11, 1, 1, 1, 3, 4, 4, 4, 3, 3, 2, 2)), comment = "#")

colnames(darter_data) <- c("Sample", "Species", "Riffle", "Age", "Sex", "Length", "Depth", "Velocity", "Temperature", "MajSub Proportion", "MinSub Proportion", "Major Type", "Minor Type")
darter_data$Riffle <- factor(darter_data$Riffle,levels=c("1", "2"))
darter_data$"Major Type" <- factor(darter_data$"Major Type",levels=c("s","fg", "sg", "lg", "c"), order = TRUE)
```

## Filtering out zonale and tetrazonum spcecies, removing MinSub and Min Type columns, renaming the MajSub Proportion and Major Type columns, aranging by Sample
```{r}
darter_plots <- darter_data %>% filter(Species != "zonale" & Species != "tetrazonum") %>% 
  select(!"MinSub Proportion" & ! "Minor Type") %>% 
  rename("major_substarte" = "MajSub Proportion", "major_type" = "Major Type") %>% 
  arrange("Sample")
```

##Plot of Legth vs. Depth of each riffle for each species
```{r}
ggplot(data = darter_plots,mapping = aes(x = Length, y = Depth, color = Species, shape = Species)) +
  geom_point() +
  facet_wrap(~Riffle, ncol = 2) +
  geom_jitter()
```
In Riffle 1, there aren't as many blennioides as in Riffle 2 and only one deeper than 30 cm.  Riffle 2 has no darters deeper than 50 cm, mostly no deeper than 30 cm, only blennioides past 22 cm.

## Box plot of Length of each riffle for each species.
```{r}
ggplot(data = darter_plots,mapping = aes(x = Length, y = Species, color = Species)) +
  geom_boxplot() +
  facet_wrap(~Riffle, ncol = 2)
```

Riffle 1 shows one outlier, Riffle 2 has the most.

## Box plot of Length of darters for each major type by Species and Riffle
```{r}
ggplot(data = darter_plots, mapping = aes(x = Length, y = major_type, color = Species))+
  geom_boxplot() +
  facet_grid(Species ~ Riffle)
```

## Same plot as above just with facet_grid variables switched.
```{r}
ggplot(data = darter_plots, mapping = aes(x = Length, y = major_type, color = Species))+
  geom_boxplot() +
  facet_grid(Riffle ~ Species)
```
The axes switch between (Species ~ Riffle) and (Riffle ~ Species)
