---
title: "HW 07: Data Importing and Wrangling 1"
author: "Ryan McAnulty"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

#### Limpets
## Loading tidyverse and reading limpets.csv data
```{r}
library(tidyverse)
read_csv("limpets.csv")
```

The data is tidy.

## Defining limpets_data and forcing column types
```{r}
limpets_data <- read_csv("limpets.csv", col_types = "ncc")
```

## Setting levels for Site 
```{r}
limpets_data$Site <- factor(limpets_data$Site, levels = c("PBL", "KNRM", "VBG", "WP", "PF", "DP", "SIO", "CTZ1", "CTZ2", "CNMM"), ordered = TRUE)
```

## Box plot of limpets_data Length (mm) vs. Site with Protected status colored
```{r}
ggplot(data = limpets_data) +
  geom_boxplot(mapping = aes(x = Length, y = Site, color = Protected)) +
  labs(x = "Length (mm)")
```
The DP and WP sites have some outliers longer than the rest of the data.  For the most part, the protected sites tend to have larger Limpets.

#### Roseate Terns
## Reading roseate_terns.txt
```{r}
read.csv("roseate_terns.txt")
```

## Naming data terns
```{r}
terns <- read.csv("roseate_terns.txt")
```

The data is not tidy

```{r}
terns %>% filter(complete.cases(.))
```

```{r}
terns %>% 
  pivot_longer(c("AR", "AL", "AC", "BR", "BL", "BC", "JR", "JL", "JC"), names_to = "sites")
```

