---
title: "HW10: Data Visualization I"
author: "Ryan McAnulty"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Loading tidyverse, here, GGally, and patchwork
```{r}
library(tidyverse)
library(here)
library(GGally)
library(patchwork)
```

## Importing sparrows.txt
```{r}
sparrow_raw <-
  read_tsv(here("data", "sparrows.txt")) # Imported using read_tsv since it is tab delineated
```

## Adding in two new columns
```{r}
sparrows <- sparrow_raw %>%
  mutate(box_group = "Sparrows",  # For arbitrary value for plotting
         row_order = 1:nrow(.))   # Setting row order from 1 to 1265
```

## Recreating Figure 2 from Zuur et al. paper
```{r}
plot_box <- sparrows %>%
  ggplot(aes(x = box_group,
             y = wingcrd)) +
  geom_boxplot(width = 0.25) +
  labs (x = NULL,
        y = "Wing length (mm)")
```

## Using a Cleveland dot plot to show full range to look for outliers
```{r}
plot_cleveland <- sparrows %>%
  ggplot(aes(x = wingcrd,
             y = row_order)) +
  geom_point() +
  labs(x = "Wing length (mm)",
       y = "Order of the data")
```

## Displaying the two plots side by side
```{r}
plot_box + plot_cleveland
```

## Flipping the Coordinates
```{r}
plot_box + plot_cleveland + coord_flip()
```
## Recreating Figure 3 of Zuur et al.
```{r}
p_culmen <-
  sparrows %>%                      # Saving culmen variable as p_culmen
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = culmen)) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
p_nalospi <- sparrows %>%           # Saving nalospi variable as p_nalospi
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = nalospi)) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
p_wt <- sparrows %>%                # Saving wt variable as p_wt
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wt)) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
p_wingcrd <- sparrows %>%           # Saving wingcrd variable as p_wingcard
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = wingcrd)) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
p_tarsus <- sparrows %>%            # Saving tarsus variable as p_tarsus
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = tarsus)) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
p_head <- sparrows %>%              # Saving head variable as p_head
  ggplot(aes(y = row_order)) +
  geom_point(aes(x = head)) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )

p_culmen + p_nalospi + p_wt + p_wingcrd + p_tarsus + p_head + plot_layout(ncol = 3) # Displaying all six life Figure 3 
```
Tarsus, head, nalospi appear to have outliers.

## Histogram on the left from Figure 5
```{r}
h1 <- sparrows %>% 
  filter(Month %in% 6:8) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 closed = "right",
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by =2))
```

## Faceted Histogram on the right from Figure 5
```{r}
h2 <- sparrows %>% 
  filter(Month %in% 6:8) %>% 
  mutate(Month = case_when(
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August")) %>% 
  ggplot() +
  geom_histogram(aes(x = wt),
                 binwidth = 0.5,
                 boundary = 0,
                 color = "black") +
  scale_x_continuous(breaks = seq(14, 28, by =2)) +
  facet_wrap(~ Month, ncol = 1)
```

## Using patchwork to assemble the two plots into a single figure
```{r}
h1 + h2
```

## Figure 10: Relationships - Using ggscatmat function from GGally package to show relationships between the data
```{r}
sparrows %>% ggscatmat(columns = 1:7)
```

## Figure 11: Interactions
```{r}
month_order = c("May", "June", "July", "August", "September")
sparrows %>% 
  filter(Month %in% 5:9,
         Sex != 0) %>% 
  mutate(Month = case_when(
    Month == 5 ~ "May",
    Month == 6 ~ "June",
    Month == 7 ~ "July",
    Month == 8 ~ "August",
    Month == 9 ~ "September"),
    Month = factor(Month, levels = month_order, ordered = TRUE),
    Sex = ifelse(Sex == 4, "Male", "Female")) %>% 
  ggplot(aes(x = wingcrd,
             y = wt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(Sex ~ Month)
```

