---
title: "HW14: Maps"
author: "Ryan McAnulty"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Loading necessary libraries
```{r}
library(here)
library(raster)
library(rgdal)
library(sf)
library(tidyverse)
library(ggthemes)
library(patchwork)
library(maptools)
library(ggmap)
library(marmap)
```

#### Part 1: Outline maps
## US Map data
```{r}
# Outline of the 48 contiguous states
usa <-map_data(map = "state") # Tidyverse or ggplot2 must be loaded

# Outline of just Missouri
us_missouri <-map_data ( map = "state",
                         region = "missouri")

# Outline of Alabama, Florida, Mississippi, and North Carolina
us_other <-subset(x = usa,
                  subset = region %in% c("alabama",
                                         "florida",
                                         "mississippi",
                                         "north carolina"))
```

## non-Missouri layer
```{r}
ggplot() +
  geom_polygon(data = us_other,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "gray")
```

## Missouri layer added
```{r}
ggplot() +
  geom_polygon(data = us_other,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "gray") +
  geom_polygon(data = us_missouri,
               aes(x = long, y = lat,
                   group = group),
               fill = "black")
```

## US Map layer added
```{r}
ggplot() +
  geom_polygon(data = us_other,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "gray") +
  geom_polygon(data = us_missouri,
               aes(x = long, y = lat,
                   group = group),
               fill = "black") +
  geom_polygon(data = usa,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               color = "black")
```

## US Map total
```{r}
# Add the filled states first so that the black outlines of all
# states are overlaid on top.
us_map <- ggplot() +
  geom_polygon(data = us_other,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = "grey") +
  geom_polygon(data = us_missouri,
               aes(x = long, y = lat,
                   group = group),
               fill = "black") +
  geom_polygon(data = usa,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               color = "black") +
  theme_map() +
  coord_fixed(1.3)
us_map
```

## Missouri
```{r}
missouri <- map_data("county", "missouri")
mo_scott <- subset(missouri, subregion %in% c("scott"))

mo_map <-ggplot() + 
  geom_polygon(data = missouri, aes(x=long, y = lat, group = group), fill = NA, color = "black") +
  geom_polygon(data = mo_scott,aes(x = long, y = lat, group = group), fill = "black") +
  theme_map() +
  coord_fixed(1.3)
mo_map
```

## US + Missouri
```{r}
us_map + mo_map + plot_layout(ncol = 2, widths = c(1.5, 1))
```

## My map US
```{r}
us_map <- ggplot() +
  geom_polygon(data = us_missouri,
               aes(x = long, y = lat,
                   group = group),
               fill = "black") +
  geom_polygon(data = usa,
               aes(x = long,
                   y = lat,
                   group = group),
               fill = NA,
               color = "black") +
  theme_map() +
  coord_fixed(1.3)
```

## My map, Missouri, Cape Girardeau, Ripley, and Dallas counties
```{r}
missouri <- map_data("county", "missouri")

mo_cape <- subset(missouri, subregion %in% c("cape girardeau"))

mo_ripley <- subset(missouri, subregion %in% c("ripley"))

mo_dallas <- subset(missouri, subregion %in% c("dallas"))

mo_map <-ggplot() + 
  geom_polygon(data = missouri, aes(x=long, y = lat, group = group), fill = NA, color = "black") +
  geom_polygon(data = mo_cape, aes(x = long, y = lat, group = group), fill = "black") +
  geom_polygon(data = mo_ripley, aes(x = long, y = lat, group = group), fill = "gray") +
  geom_polygon(data = mo_dallas, aes(x = long, y = lat, group = group), fill = "gray") +
  theme_map() +
  coord_fixed(1.3)
```

## My maps combined
```{r}
us_map + mo_map + plot_layout(ncol = 2, widths = c(1.5, 1))
```
#### Part 2: Choropleth maps
## Importing skin cancer raw data
```{r}
skin_cancer_raw <- 
  read_delim(file = here::here("data","skin_cancer.txt"),
             delim = " ",
             skip = 7,
             trim_ws = TRUE)
```

## Wrangling the data
```{r}
skin_cancer <- skin_cancer_raw %>% 
  filter(State != "Wash, D.C.") %>% 
  mutate(State = case_when(
    State == "NorthCarolina" ~ "North Carolina",
    State == "SouthCarolina" ~ "South Carolina",
    State == "NorthDakota" ~ "North Dakota",
    State == "SouthDakota" ~ "South Dakota",
    State == "NewHampshire" ~ "New Hampshire",
    State == "NewJersey" ~ "Nes Jersey",
    State == "NewMexico" ~ "New Mexico",
    State == "MewYork" ~ "New York", # Data has MewYork typo
    State == "RhodeIsland" ~ "Rhode Island",
    State == "WestVirginia" ~ "West Virginia",
    TRUE ~ State
  ))
```

## Mean Mortality
```{r}
mean_mort <-mean(skin_cancer$Mort)

skin_cancer_adj <- skin_cancer %>%
  mutate(relative_mort = Mort-mean_mort)
```

## Lower 48
```{r}
# Used to trim Alaska, Hawaii, and territories from the shapefile.
lower_48 <-c("Alabama","Arizona","Arkansas","California","Colorado","Connecticut","Delaware","Florida","Georgia","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey","New Mexico","New York","North Carolina","North Dakota","Ohio","Oklahoma","Oregon","Pennsylvania","Rhode Island","South Carolina","South Dakota","Tennessee","Texas","Utah","Vermont","Virginia","Washington","West Virginia","Wisconsin","Wyoming")
```

## States
```{r}
states <-st_read(here::here("data", "cb_2017_us_state_500k.shp"),
                 stringsAsFactors = TRUE)
```

## States_df
```{r}
states_df <- states %>% 
  dplyr::filter(NAME %in% lower_48)
```

## FIPS factor
```{r}
str(states_df$STATEFP)
```
##
```{r}
# Capture the FIPS levels from states_df
# Use to factor the FIPS and state names in the next step.
state_levels <-levels(states_df$STATEFP)
name_levels <-levels(states_df$NAME)
```

## fips again
```{r}
fips <-read_csv(here::here("data", "state_fips.csv")) %>% 
  select(-abbr) %>%
  rename(STATEFP = FIPS) %>%
  mutate(STATEFP =factor(STATEFP, levels = state_levels))
```

## Left Joins
```{r}
skin_cancer_adj <-left_join(skin_cancer_adj, fips)
states_df <-left_join(states_df, skin_cancer_adj)
```
## Skin Cancer Mortality Map
```{r}
ggplot(states_df) +
  geom_sf(aes(fill = relative_mort)) +
  scale_fill_viridis_c(name = "Skin Cancer,\nRelative Mortality",
                       option = "inferno") +
  coord_sf(crs =st_crs(5070)) +
  theme_bw() +
  theme(legend.position = "bottom")
```

## My Choropleth map
```{r}
lyme_disease_raw <- 
  read_csv(file = here::here("data","lyme_disease.csv"),
             skip = 1,
             na = c("-"))

lyme_disease <- lyme_disease_raw %>% 
  mutate(`2017` = `2017_Confirmed` + `2017_Probable`) %>% 
  select(State:`2016`, `2017`) %>% 
  pivot_longer("2007":"2017",
               names_to = "Year",
               values_to = "cases") %>% 
  group_by(State) %>% 
  summarise(mean_cases = mean(cases),
            log_mean = log10(mean_cases),
            .groups = "drop")

  lyme_disease
```

## Merging data
```{r}
lyme_disease <-left_join(lyme_disease, fips)
states_df <-left_join(states_df, lyme_disease)
```
## Plot of my map
```{r}
ggplot(states_df) +
  geom_sf(aes(fill = log_mean)) +
  scale_fill_viridis_c(name = "Lyme Disease Mean \nCases by State",
                       option = "cividis", labels = c("<1", "10", "100", "1000", "5000")) +
  coord_sf(crs =st_crs(5070)) +
  theme_bw() +
  theme(legend.position = "bottom")
```

#### Part 3: Dot distribution maps
## Distribution of Gigantopithecus and Ursus
```{r}
# Global vars -------------------------------------------------------------# Restrict the map to the western US.
min_long <- -126
max_long <- -104
min_lat  <-   31
max_lat  <-   50

set.seed(9618973)
```

## Bigfoot data
```{r}
# Bigfoot data ------------------------------------------------------------
bigfoot <-read_csv(here::here("data", "bigfoot.csv"))
bigfoot <- bigfoot %>%
  filter(long >= min_long & long<=max_long, 
         lat>=min_lat & lat<=max_lat) %>%
  sample_n(300) %>%
  mutate(name = "Gigantopithecus") %>%
  dplyr::select(name, long, lat)
```

## Bear data
```{r}
# Bear data ---------------------------------------------------------------
bears <- read_csv(here::here("data", "bears.csv"))

both_species <- bind_rows(bigfoot, bears)

head(both_species)
```
```{r}
tail(both_species)
```
## Stamen Maps
```{r}
base = get_stamenmap(bbox = c(min_long,
                              min_lat,
                              max_long,
                              max_lat),
                     zoom = 4,
                     maptype = "terrain-background")

ggmap(base)
```

## Bigfoot Map
```{r}
base <-ggmap(base)

base + geom_point(data = bigfoot,
                  aes(x = long,
                      y = lat))
```

## Bears Map
```{r}
base + geom_point(data = bears,
                  aes(x = long,
                      y = lat))
```

## Both Species
```{r}
base + geom_point(data = both_species,
                  aes(x = long,
                      y = lat,
                      shape = name,
                      fill = name),
                  size = 2) + 
  scale_shape_manual(values =c(22:23)) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Species",
       shape = "Species")
```

## Distribution map: my map
```{r}
# Global vars -------------------------------------------------------------# Restrict the map to the western US.
min_long <- -126
max_long <- -104
min_lat  <-   31
max_lat  <-   50

set.seed(3926003)
```

## Bigfoot data
```{r}
# Bigfoot data ------------------------------------------------------------
bigfoot <-read_csv(here::here("data", "bigfoot.csv"))
bigfoot <- bigfoot %>%
  filter(long >= min_long & long<=max_long, 
         lat>=min_lat & lat<=max_lat) %>%
  sample_n(300) %>%
  mutate(name = "Gigantopithecus") %>%
  dplyr::select(name, long, lat)
```

## Bear data
```{r}
# Bear data ---------------------------------------------------------------
bears <- read_csv(here::here("data", "bears.csv"))

both_species <- bind_rows(bigfoot, bears)

head(both_species)
```
```{r}
tail(both_species)
```
## Stamen Maps
```{r}
base = get_stamenmap(bbox = c(min_long,
                              min_lat,
                              max_long,
                              max_lat),
                     zoom = 4,
                     maptype = "terrain-background")

ggmap(base)
```

## Bigfoot Map
```{r}
base <-ggmap(base)

base + geom_point(data = bigfoot,
                  aes(x = long,
                      y = lat))
```

## Bears Map
```{r}
base + geom_point(data = bears,
                  aes(x = long,
                      y = lat))
```

## Both Species
```{r}

base + geom_point(data = both_species,
                  aes(x = long,
                      y = lat,
                      shape = name,
                      fill = name),
                  size = 2) + 
  scale_shape_manual(values =c(22:23)) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = "Species",
       shape = "Species")
```
#### Part 4: Bathymetry:  Due to Mr. Gates' contributions to the computing world.  Unable to complete.

