---
title: "HW13: Strings and Things"
author: "Ryan McAnulty"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Loading required libraries
```{r}
library(tidyverse)
library(here)
library(ggthemes)
library(gghighlight)
```
## From many strings to one
```{r}
beta <-                                   # Naming imported data
  read_lines(here("data",                 # Importing "beta_gliadin.txt"   
                  "beta_gliadin.txt"),
                  skip = 1)               # Skipping meta data

DNA <- beta %>%                           # Saving changes
  str_replace_all("[:digit:]",            # Removing numerals and replacing with nothing
                  "") %>%         
  str_replace_all("[:space:]",            # Removing white spaces and replacing with nothing
                  "") %>%          
  str_c(collapse = "") %>%                # Collapsing 56 strings into one
  str_to_upper()                          # Converting to all Upper case characters

DNA                                       # Displaying string to prove changes worked
str_count(DNA)
```
There are 3310 characters
## Sequence analysis: find the promoter region
## Counting non ACTG
```{r}
str_count(DNA,
          "[^ACTG]")   # REGEX anything but
```
There are 0 non A, C, T, or G characters.

# Counting starting codon ATG
```{r}
str_count(DNA,
          "ATG")
```
66 ATGs.

## Counting stop codon TGA or TAA
```{r}
str_count(DNA, 
          "TGA|TAA")  # or
```
99 TGAs or TAAs.

## Finding Promoter region
## Assigning "CAAT box" and "TATA box"
```{r}
caat_box <- "CCAAT"
tata_box <- "CTATAA"
```

## Counting strings
```{r}
str_count(DNA,  
          caat_box)                    # There are 4 instances of "CCAAT" in the string
str_count(DNA, 
          tata_box)                    # There is 1 instance of #CTATAA" in the string 

caat_seq <- "CCAAT.{100,150}ATG"       # Making a sequence search that searches for a sequence that starts with "CCAAT" and ends with the stop codon "ATG" with 100 to 150 nucleotides in between
str_count(DNA, 
          caat_seq)                    # There are 2 such sequences
tata_seq <- "CTATAA.{50,100}ATG"       # Making a sequence search that searches for a sequence that starts with "CTATAA" and ends with the stop codon "ATG" with 50 to 100 nucleotides
str_count(DNA, 
          tata_seq)                    # Just 1 of these
caat_seqs_all <- 
  str_extract_all(DNA,                 # Extracting the two "CTATAA" sequences
                  caat_seq, 
                  simplify = TRUE)
str_length(caat_seqs_all)              # Lengths are 139 and 148 respectivley for these two
str_which(caat_seqs_all, tata_box)     # Searching for sequence that contains the "TATA box"
promoter <- str_sub(caat_seqs_all[1])  # Naming this promoter
promoter                               # Displaying promoter
str_length(promoter)                   # Sequence length is the shorter 139 length seq
```
#### Part 2: Ants
## Import the data
```{r}
ants<- read_csv(here("Data", "ants.csv"))
ants
```

## Wrangle the data
## Species and abundance
```{r}
ants_species <- ants %>% 
  pivot_longer(names_to = "species",                     # pivoting data to spcecies and abundance 
               values_to = "abundance", 
               matches("[[:upper:]].+\\.[[:lower:]].+",  # Grabbing capitalized Genus names seperated by a . then species   
                       ignore.case = FALSE)) %>%         # Don't ignore case
  select(year,                                           # Only displaying year, species, and abundance  
         species, 
         abundance)
```

## Creating summary of species, converting species to CEP names, with total abundance
```{r}
ants_sum <- ants_species %>% 
  mutate(species = 
           str_replace(species,         # Changing species names to CEP names
                             "(^[A-Z][a-z]{0,3}).*\\.([a-z]{1,4}).*",
                             "\\1\\2")) %>% 
  mutate(year = factor(year,            # Creating an ordered year factor
                ordered = TRUE,
                levels = c(2010,        # levels of 2010 and 2011
                           2011))) %>% 
  group_by(year,                        # Grouping by year and species 
           species) %>% 
  summarise(total_abundance =           # Summarizing with total abundance
              sum(abundance),           # Adding abundances together
            .groups = "drop")           # Dropping groups
ants_sum
```
# Cleveland plot of ants summary
```{r}
ants_sum %>% 
  ggplot(mapping = 
           aes(x = reorder(species,            # Reordering species and total_abundance and setting it to x
                           total_abundance), 
               y = total_abundance,            # Setting total_abundance to y axis   
               group = species)) +             # grouping abundance
  geom_line() +                                # Line graph layer 
  geom_point(aes(shape =                       # Point layer setting years as different shapes at size 2
                   year),
             size = 2) +
  coord_flip() +                               # Flipping the coordinates
  labs(x = NULL,                               # Setting x-axis as NULL, y as Total Abundance  
       y = "Total Abundance", 
       shape = "Year") +                       # Changing plot title to "Year"
  theme_minimal()                              # Simplifies plot
```

## Code for slope graph
```{r}
ants_sum %>%
  filter(total_abundance>=30) %>% # Arbitrary value
  ggplot(aes(x = year, y = total_abundance, group = species)) +
  geom_line(color = "gray") +
  geom_point() +
  theme_classic() + 
  geom_text(data = ants_sum %>%
              filter(year == "2010",
                     total_abundance >= 30),
            aes(x = year,
                y = total_abundance,
                label = species,
                fontface = "italic"),
            hjust = 1.1) +
  geom_text(data = ants_sum %>% 
            filter(year == "2011",
                      total_abundance >= 30),
            aes(x = year,
                y = total_abundance,
                label = species,
                fontface = "italic"),
            hjust = -0.1) +
  labs(x = NULL,
       y = "Abundance") +
  theme(legend.position = "none") + 
  scale_x_discrete(expand =c(0.2, 0.02))
```
Almost every ant species decreased from 2011 to 2010 due to the burning.
Floreprui went from the most populated down to the five others with more than 30 of abundance.

#### Part 3: Featuring Phlox Flowers
## Import and wrangle data
```{r}
flower_size <- read.csv(here("data",
                             "flower_size.csv"))  # Importing Flower_size data   
  
flower_size_summary <- flower_size %>%
  drop_na(Flower.number) %>%                      # Dropping nas in column Flower.number
  select("Species",                               # Selecting "Species", "cor_length", "throat_width", renaming the latter two
         "cor_length" = 
           Corolla.length..cm.,
         "throat_width" = 
           Corolla.width.throat..cm.) %>% 
  separate(Species,                               # Separating the species column into genus and species  
           into = c("genus", 
                    "species"), 
           sep = " ") %>%                         # Denoting the deliminator is a space
  group_by(genus, species) %>%                    # Grouping by genus and species
  summarise("mean_length" = mean(cor_length, na.rm = TRUE),     # Creating summary for mean length, mean width, max length and max width
            "mean_width" = mean(throat_width, na.rm = TRUE),
            "max_length" = max(cor_length, na.rm = TRUE),
            "max_width" = max(throat_width, na.rm = TRUE),      
  .groups = "drop")                               # Dropping groups 

flower_size_summary
```

## Importing and wrangling the second data set
```{r}
pollinators_raw <- read.csv(here("data", 
                                 "flower_pollinators.csv"))  # Importing data

pollinators <- pollinators_raw %>%                 # Naming data
  select(-(starts_with("Source") | Color)) %>%     # Removing "Source" and Color columns
  filter(Pollinator != "") %>%                     # Filtering out NAs
  separate(ï..Species,                             # Separating species into genus and species
            into = c("genus",
                     "species"),
            sep = " ") %>%  
  separate(Pollinator, 
           into = "Syndrome", sep = "(,|) ")
pollinators
```
## Using inner_join to combine data sets
```{r}
phlox <- inner_join(flower_size_summary,
                    pollinators)
phlox
```

## Plotting Part 3
```{r}
phlox %>% 
  ggplot(mapping = aes(x = mean_length, 
                       y = mean_width, 
                       color = Syndrome)) +    # Scatter plot of phlox
           geom_point() +
           labs(x = "Mean length",            # Renaming axes 
                y = "Mean width")
```
## Fixing "bee" and "beetle
```{r}
p_1 <- phlox %>% 
  mutate(Syndrome = 
           str_replace(Syndrome,           # Search in Syndrome for pollinator that begins with be(any character) ends with e
                       "(be.*e$)", 
                       "\\1s")) 
p_1 %>% 
  ggplot(mapping = aes(x = 
                         mean_length, 
                       y = 
                         mean_width, 
                       color = Syndrome))+    # Scatter plot of phlox
           geom_point() +
           labs(x = "Mean length",            # Renaming axes 
                y = "Mean width")
```
## Looking at low pollinator counts
```{r}
p_1 %>% 
  group_by(Syndrome) %>% 
  count(Syndrome)
```
## Creating other insects out of beetles and flies
```{r}
p_2 <- p_1 %>% 
  mutate(Syndrome =
           str_replace(Syndrome, 
                       ".+l.?es", 
                       "other insects"))
p_2 %>% 
  ggplot(mapping = aes(x = 
                         mean_length, 
                       y = 
                         mean_width, 
                       color = Syndrome))+    # Scatter plot of phlox
           geom_point() +
           labs(x = "Mean length",            # Renaming axes 
                y = "Mean width")
```


## Making Mammal, grouping insects, capitalizing autogamous
```{r}
p_3 <- p_2 %>% 
  mutate(syndrome_group = case_when(
           Syndrome == "bat" ~ "Mammal",
           Syndrome == "hummingbird" ~ "Bird",
           Syndrome == "autogamous" ~ "Autogamous",
           TRUE ~ "Insect"))

p_3 %>%
    ggplot(mapping = aes(x = 
                         mean_length, 
                       y = 
                         mean_width, 
                       color = syndrome_group))+    # Scatter plot of phlox
           geom_point() +
           labs(x = "Mean length",            # Renaming axes 
                y = "Mean width",
                color = "Syndrome")
  
 
```

## gghighlight
```{r}
p_3 %>%
    ggplot(mapping = aes(x = 
                         mean_length, 
                       y = 
                         mean_width, 
                       color = syndrome_group))+    # Scatter plot of phlox
           geom_point() +
           labs(x = "Mean length",                  # Renaming axes 
                y = "Mean width",
                color = "Syndrome") +
           gghighlight(syndrome_group ==            # Using gghighlight to highlight Mammal and Bird
                         "Mammal" | 
                         syndrome_group == 
                         "Bird", 
                       use_direct_label = FALSE)    # Keeping Legend
```

## Colors and themes
```{r}
p_3 %>%
  ggplot()+
  geom_point(aes(x = max_length,y = max_width,color = syndrome_group))+ #  scale_color_viridis_d(option = "viridis")
  scale_color_brewer(type = "qual", 
                     palette = "Dark2") +
  labs(x = "Mean Length (cm)",y = "Mean Width (cm)",color = "Syndrome")
```
## Semo pallete
```{r}
# Custom color palette, using
# Southeast Red, Gum Tree, Fountain, Copper
semo_palette <-c("#C8102E", "#DBE245", "#71C5E8", "#B07C57")

p_3 %>%
  ggplot() +
  geom_point(aes(x = mean_length,
                 y = mean_width,
                 shape = syndrome_group,
                 fill = syndrome_group)) +
  scale_shape_manual(values =c(22:25)) +
  #  scale_fill_brewer(palette = "Dark2")
  #  scale_fill_viridis_d(option = "viridis")
  scale_fill_manual(values = semo_palette) +
  labs(x = "Mean Length (cm)",
       y = "Mean Width (cm)",
       shape = "Syndrome",
       fill = "Syndrome")
```
## ggthemes
```{r}
p_3 %>%
  ggplot() +
  geom_point(aes(x = mean_length,
                 y = mean_width,
                 shape = syndrome_group,
                 fill = syndrome_group),
             size = 2) +
  labs(shape = "Syndrome",
       fill = "Syndrome",
       x = "Mean Length (cm)",
       y = "Mean Width (cm)") +
  theme_tufte() +
  scale_shape_manual(values = c(22:25)) + 
  scale_fill_brewer(palette = "Dark2") +
  theme(axis.text = element_text(size = 10),
        axis.title =element_text(size = 12))
```
## St. Louis Cadinals color palette 
```{r}
# Custom color palette, using
# Navy, Cardinal Red, Yellow, White
Cardinal_palette <-c("#0C2340", "#C41E3A", "#FEDB00", "#FFFFFF")

p_3 %>%
  ggplot() +
  geom_point(aes(x = max_length,
                 y = max_width,
                 shape = syndrome_group,
                 fill = syndrome_group)) +
  scale_shape_manual(values =c(22:25)) +
  #  scale_fill_viridis_d(option = "viridis")
  scale_fill_manual(values = Cardinal_palette) +
  labs(x = "Max Length (cm)",
       y = "Max Width (cm)",
       shape = "Syndrome",
       fill = "Syndrome") + 
  scale_shape_manual(values = c(22:25)) + 
  theme(axis.text = element_text(size = 14),
        axis.title =element_text(size = 14))
```
## Fallout New Vegas Color palette 
```{r}
# Custom color palette, using
# darkest night, Round Tuit, Sticky Wicket, Mummy Glow
New_vegas_palette <-c("#331111", "#665533", "#995533", "#DDCC66")

p_3 %>%
  ggplot() +
  geom_point(aes(x = max_length,
                 y = max_width,
                 shape = syndrome_group,
                 fill = syndrome_group)) +
  scale_shape_manual(values =c(22:25)) +
  scale_fill_manual(values = New_vegas_palette) +
  labs(x = "Max Length (cm)",
       y = "Max Width (cm)",
       shape = "Syndrome",
       fill = "Syndrome")
```

## Ugly plot
```{r}
p_3 %>%
  ggplot() +
  geom_point(aes(x = max_length,
                 y = max_width,
                 shape = syndrome_group,
                 fill = syndrome_group),
             size = 10) +
  labs(shape = "Syndrome",
       fill = "Syndrome",
       x = "Max Length (cm)",
       y = "Max Width (cm)") +
  theme_tufte() +
  scale_shape_manual(values = c(20:23)) + 
  scale_fill_brewer(palette = "PiYG") +
  theme(axis.text = element_text(size = 10),
        axis.title =element_text(size = 12))
```

## Really good looking figure
```{r}
# Custom color palette, using
# Red, Yellow, Green, Purple
Rainbow_palette <-c("#ff0000", "#ffff00", "#008000", "#4b0082")

p_3 %>%
  ggplot() +
  geom_point(aes(x = max_length,
                 y = max_width,
                 shape = syndrome_group,
                 fill = syndrome_group),
             size = 4) +
  scale_shape_manual(values =c(21:24)) +
  scale_fill_manual(values = Rainbow_palette) +
  labs(x = "Max Length (cm)",
       y = "Max Width (cm)",
       shape = "Syndrome",
       fill = "Syndrome") +
      theme(axis.text = element_text(size = 10),
        axis.title =element_text(size = 10))
```

